---
title: 'Homework #3'
author: "Guang Yi Lim;gylim0604"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Tips at Restaurants

In this homework we will use the `tips` data set. This data set is part of the `reshape2` package. You can load the data set by executing the command:
```{r}
data(tips, package="reshape2")
library(ggplot2)
```
The information contained in the data is collected by one waiter, who recorded over the course of a season information about each tip he received working in one restaurant. See `?tips` for a description of all of the variables. 

**Submission instructions**: Create a folder named `ds202_hw3`, and name the RMarkDown file including your solutions `hw3.Rmd` under this folder. For submission, create a GitHub repository named `ds202_hw3` under your GitHub account, and push both `hw3.Rmd` and the knitted `hw3.html` before the deadline. I will assume you use the same GitHub username as for your HW2 submission. The latest version of your homework appearing on GitHub before the deadline is graded. *It is crucial to follow the exact folder structure and file names*, so that your homework can be reproduced and graded by an automated script. This homework is **due on Feb 19** before class.

1.  How many parties did the waiter serve? Store the number in `numParty` and print. 
```{r}
numParty<-nrow(tips)
numParty
```
</br>
2. What are the types of variables in the data frame `tips`? 
```{r}
str(tips)
```
</br>
3. Create a vector named `day` containing the day of visits in `tips`. The factor levels should be ordered from Thursday to Sunday. Also print the variable.
```{r}
day <- factor(tips$day,levels = c("Thur","Fri","Sat","Sun"))
day
```
</br>
4. Create a data frame named `female5` containing the meal paid by a female payer in a  party with size greater than or euqal to 5. Also print the data frame.
```{r}
female5 <-  subset(tips, sex== "Female" & size >=5)
female5
```
</br>
5. How does the tipping amount (`tip`) depend on the overall bill (`total_bill`)? Describe the relationship in words.
Describe at least two types of anomalies in the plot. What do they mean? Use the `ggplot2` package to make a chart.
```{r}
#tipping increases as the total bill. In the graph, there was a large top amount for a small total bill. the tipping amount between 40 to 45 are lower than expected.
ggplot(tips,aes(x=total_bill,y=tip)) +geom_point()
```
</br>
6. Introduce a variable `tiprate` into the data set, that incorporates the rate of tips. What is the average rate for tips?
```{r}
tips$tiprate = tips$tip/tips$total_bill
mean(tips$tiprate)
```
</br>
7. Make a boxplot of the tip rate by time. The x-axis should be ordered by lunch and then dinner. Use the `ggplot2` package to make a chart.
```{r}
temp <- factor(tips$time,levels =c("Lunch","Dinner"))
ggplot(tips,aes(x= temp, y =tiprate))+geom_boxplot()
```
</br>
8. How does smoking behavior and gender of the person who pays impact the  relationship between tip and total bill? Find a visualization that incorporates all four variables. Interpret the result.
```{r}
temp <-paste(tips$sex,tips$smoker)
df <-tips
df$temp <-temp
ggplot(df,aes(x =total_bill,y=tip,color = temp)) +geom_point()

#Males pay higher total bill compared to females and they also tip higher compared to females.
#males who dont smoke tend to tip higher than those who smoke
#on average females who smoke tips slightly lower than females who dont
```
</br>
9. Use ggplot2 to find a graphical summary of the relationship between day of the week and gender of the person paying the bill. What can you say about this relationship?
```{r}
ggplot(tips,aes(x=factor(day,levels = c("Thur","Fri","Sat","Sun")),fill=sex))+geom_bar()
#On weekdays, the ratio between male to female who pay for meals are almost the same. However, Males pay more often during weekends
```

Note: your submission is supposed to be fully reproducible, i.e. the TA and I will 'knit' your submission in RStudio. 

